\vspace{-0.8in}Brief Announcement: A Family of Leaderless Generalized-Consensus Algorithms

### 2016-02-12 podc16/paper2.tex
Working on the implementation section.

We consider a set of processes $P$ subject to crash-stop faults and communicating by message-passing in an asynchronous network. Processes in $P$ must solve the Generalized Consensus problem~\cite{Lamport05GeneralizeConsensus}. Each process receives proposals for commands of the form $\opn{Propose}(c)$, and must produce responses of the form $\opn{Decide}(\sigma)$, where $\sigma$ is a c-struct. 
The sequence of calls observed must satisfy the non-triviality, consistency, stability, and liveness properties of generalized consensus.
We use the same notation and refer the reader to~\cite{Lamport05GeneralizeConsensus} for a thourough discussion of c-structs and generalized consensus. Briefly, the c-structs are a set, including $\bot$, with an operator $\bullet$ for appending commands to a c-struct. Intuitively, a c-struct represents a set of sequences that are all equivalent up to the ordering of commutative commands.
We say that two commands \textit{commute} when $\sigma\bullet c_1\bullet c_2 =  \sigma\bullet c_2\bullet c_1 $, for every c-struct $\sigma$.
C-structs are partially ordered: $\sigma_1\sqsubseteq \sigma_2$ iff there exists a sequence of commands $\mathcal{D}$ such that $\sigma_2 = \sigma_1\bullet \mathcal{D}$ and $\bot\sqsubseteq \sigma$ for any
$\sigma$. Two c-structs are compatible when they have a
common upper bound. A c-struct contains a command $c$ when it is of the form $\bot\bullet cs$ where $c$ appears in the sequence of commands $cs$. The consistency property of generalized consensus requires that any two decided c-structs be compatible.

Our leaderless generalized consensus algorithm orchestrates a dependency-set algorithm and map-agreement algorithm as shown in \cref{fig:deps-set-algo3}.
A process $p$ maintains two local variables $m_p$ (a map from commands to sets of commands) and $\sigma_p$ (a c-struct). Upon  receiving a command $c$, $p$ proceeds in three phases. First, it calls a \textit{dependency-set} algorithm (\Cref{dep-algo}) to determine a set of commands $\mathcal{D}$ that the command $c$ may depend on. Second, $p$ proposes the mapping  $c\mapsto \mathcal{D}$ to a \textit{map-agreement} algorithm (\Cref{map-algo}); the map-agreement algorithm may subsequently commit a mapping $c\mapsto \mathcal{D}'$ for $c$, in which case $p$ inserts the mapping $\left[c \mapsto \mathcal{D}'\right]$ in $m_p$, and we say that $\mathcal{D}'$ is the dependency set of $c$.
We say that $c$ is executable in $m_p$ when all its transitive dependencies are known. %no command $c'$ for which no mapping has been committed is reachable from $c$ by recursively following committed mappings. 
Finally, when $c$ is executable in $m_p$, then $p$ may execute $c$ by deciding the c-struct obtained by \emph{locally} running a \emph{graph-processing} algorithm described in \Cref{abstract-algo}.
Any other process can run the three phases independently of $p$ and eventually execute the command despite a crash of $p$.
We show that the leaderless  generalized consensus algorithm guarantees the safety and liveness properties of generalized consensus when its underlying dependency-set and map-agreement algorithms satisfy their safety and liveness properties.

@@

A dependency-set algorithm exposes the input interface $\opn{announce}_p\left( c \right)$, to announce a command $c$, the output interface $\opn{suggest}_p \left( c,\mathcal{D}\right)$, to suggest a set of dependencies $\mathcal{D}$ for $c$, and an input interface $\opn{commit}\left( c,\mathcal{D} \right)$ (meant to observe the map-agreement algorithm), where $p\in P$ is the process executing the call.
Moreover, a dependency-set algorithm must ensure that:
\begin{itemize}[noitemsep,nolistsep]
    \item[-] Safety: (\setword{S1}{dep-valid}), for any suggestion $\opn{suggest}_p \left( c,\mathcal{D}\right)$, every command in $\{c\}\cup \mathcal{D}$ must have been announced before; (\setword{S2}{suggestion-agreement}), for any two suggestions $\opn{suggest}_p \left( c_1,\mathcal{D}_1\right)$ and $\opn{suggest}_q \left( c_2,\mathcal{D}_2\right)$, if $c_1$ and $c_2$ do not commute then either $c_1\in \mathcal{D}_2$ or $c_2 \in \mathcal{D}_1$.
%    \item If $c$ was announced by $p$, then eventually a dependency set is suggested for $c$ to $p$.
    \item[-] Liveness: (\setword{L1}{eventual-suggestion}) If a suggestion $\opn{suggest}_p\left( c \right)$ was made, then eventually a suggestion $\opn{suggest}\left( c,\mathcal{D} \right)_p$ is made, and (\setword{L2}{re-suggest}), if a suggestion $\opn{suggest}_p(c,\mathcal{D})$ has been made and no input $\opn{commit}_q\left( c,\mathcal{D'} \right)$ is observed, then eventually a new suggestion $\opn{suggest}_r\left( c,\mathcal{D''} \right)$ for $r\neq p$ is made.
\end{itemize}

An example of implementation of such an algorithm is depicted by the pseudocode of \cref{fig:deps-set-algo1,fig:deps-set-algo2}. Due to space constraints, our example focuses on satisfying the safety property only. The main idea is the following: a process $p$ announcing a command $c$ broadcasts $c$ to all the processes, and it suggests the dependency-set $deps_p[c]$ for $c$ as the union of the sets $\mathcal{D}$ that it receives as replies from a quorum of $\mathcal{Q}$ processes; on the other hand, a process that receives an announcement for $c$, computes its reply $\mathcal{D}$ for $c$ as the set of commands $d$ that it received so far, and such that $c$ and $d$ do not commute. The choice of the value of $\mathcal{Q}$ must be such that the intersection of any two quorums is not empty. About correctness, we can easily prove that such an implementation satisfies the safety property by noticing that:  any process $p$ inserts a command $d$ in a dependency-set $\mathcal{D}$ that $p$ sent as a reply for a command $c$, only if $p$ previously received $d$. Therefore, every command suggested in a dependency-set must have been announced before. Further, since a process can suggest a set of dependencies $deps_p[c]$ for a command $c$ after it has received a reply for $c$ from at least a quorum of processes, and thanks to the requirement that we enforced on the intersection of quorums, we can conclude that for any two suggestions $\opn{suggest}_p \left( c_1,deps_p[c_1]\right)$ and $\opn{suggest}_q \left( c_2,deps_q[c_2]\right)$,  if $c_1$ and $c_2$ do not commute then either $c_1\in deps_q[c_2]$ or $c_2 \in deps_p[c_1]$.

@@

An example of implementation of the map-agreement algorithm can be any state machine replication algorithm, like MultiPaxos~\cite{lamport2001paxos}, where a pair of command $c$ and dependency-set $\mathcal{D}$ can be proposed to a instance of MultiPaxos that is identified by the command $c$ itself.
%used to implementing map-agreement by uniquely associating commands to MultiPaxos instance numbers.

\subsection{Processing the Dependency Graph}
\label{abstract-algo}

@@

\subsection{Obtaining EPaxos}

\printbibliography%

\appendix

@@

\end{document}

>>>>

We consider a set of processes $P$ subject to crash-stop faults and communicating by message-passing in an asynchronous network. Processes in $P$ must solve the Generalized Consensus problem~\cite{Lamport05GeneralizeConsensus}. Each process receives proposals for commands of the form $\opn{Propose}(c)$, and must produce responses of the form $\opn{Decide}(\sigma)$, where $\sigma$ is a c-struct. 
The sequence of calls observed must satisfy the non-triviality, consistency, stability, and liveness properties of generalized consensus.
We use the same notation and refer the reader to~\cite{Lamport05GeneralizeConsensus} for a thourough discussion of c-structs and generalized consensus. Briefly, the c-structs are a set, including $\bot$, with an operator $\bullet$ for appending commands to a c-struct. Intuitively, a c-struct represents a set of sequences that are all equivalent up to the ordering of commutative commands.
We say that two commands \textit{commute} when $\sigma\bullet c_1\bullet c_2 =  \sigma\bullet c_2\bullet c_1 $, for every c-struct $\sigma$.
C-structs are partially ordered: $\sigma_1\sqsubseteq \sigma_2$ iff there exists a sequence of commands $s$ such that $\sigma_2 = \sigma_1\bullet s$ and $\bot\sqsubseteq \sigma$ for any
$\sigma$. Two c-structs are compatible when they have a
common upper bound. A c-struct contains a command $c$ when it is of the form $\bot\bullet cs$ where $c$ appears in the sequence of commands $cs$. The consistency property of generalized consensus requires that any two decided c-structs be compatible.

Our leaderless generalized consensus algorithm orchestrates a dependency-set algorithm and map-agreement algorithm as shown in \cref{fig:deps-set-algo3}.
A process $p$ maintains two local variables $m_p$ (a map from commands to sets of commands) and $\sigma_p$ (a c-struct). Upon  receiving a command $c$, $p$ proceeds in three phases. First, it calls a \textit{dependency-set} algorithm (\Cref{dep-algo}) to determine a set of commands $\mathcal{D}$ that the command $c$ may depend on. Second, $p$ proposes the mapping  $c\mapsto \mathcal{D}$ to a \textit{map-agreement} algorithm (\Cref{map-algo}); when the map-agreement algorithm
commits a mapping $c\mapsto \mathcal{D}'$ for $c$, $p$ inserts the mapping $\left[c \mapsto \mathcal{D}'\right]$ in $m_p$, and we say that $\mathcal{D}'$ is the dependency set of $c$.
We say that $c$ is executable in $m_p$ when all its transitive dependencies are known. %no command $c'$ for which no mapping has been committed is reachable from $c$ by recursively following committed mappings. 
Finally, when $c$ is executable in $m_p$, then $p$ may execute $c$ by deciding the c-struct obtained by \emph{locally} running a \emph{graph-processing} algorithm described in \Cref{abstract-algo}.
Any other process can run the three phases independently of $p$ and eventually execute the command, possibly despite a crash of $p$.
%We show that the leaderless  generalized consensus algorithm guarantees the safety and liveness properties of generalized consensus when its underlying dependency-set and map-agreement algorithms satisfy their safety and liveness properties.

@@

A dependency-set algorithm exposes at each process $p$ the input interface $\opn{announce}_p\left( c \right)$, to announce a command $c$, the output interface $\opn{suggest}_p \left( c,\mathcal{D}\right)$, to suggest a set of dependencies $\mathcal{D}$ for $c$, and an input interface $\opn{commit}\left( c,\mathcal{D} \right)$,  to observe which command is committed by the other abstraction.
Moreover, a dependency-set algorithm must ensure that:
\begin{itemize}[noitemsep,nolistsep]
    \item[-] Safety: (\setword{S1}{dep-valid}), for any suggestion $\opn{suggest}_p \left( c,\mathcal{D}\right)$, every command in $\{c\}\cup \mathcal{D}$ must have been announced before; (\setword{S2}{suggestion-agreement}), for any two suggestions $\opn{suggest}_p \left( c_1,\mathcal{D}_1\right)$ and $\opn{suggest}_q \left( c_2,\mathcal{D}_2\right)$, if $c_1$ and $c_2$ do not commute then either $c_1\in \mathcal{D}_2$ or $c_2 \in \mathcal{D}_1$.
%    \item If $c$ was announced by $p$, then eventually a dependency set is suggested for $c$ to $p$.
    \item[-] Liveness: (\setword{L1}{eventual-suggestion}) If a suggestion $\opn{suggest}_p\left( c \right)$ was made, then eventually a suggestion $\opn{suggest}\left( c,\mathcal{D} \right)_p$ is made, and (\setword{L2}{re-suggest}), if a suggestion $\opn{suggest}_p(c,\mathcal{D})$ has been made and no input $\opn{commit}_q\left( c,\mathcal{D'} \right)$ is observed, then eventually a new suggestion $\opn{suggest}_r\left( c,\mathcal{D''} \right)$ for $r\neq p$ is made.
\end{itemize}

@@

\subsection{Local Dependency Graph Processing}
\label{abstract-algo}

@@

\subsection{Implementing the Abstractions}
Both the dependency-set and map-agreement abstractions are well suited to leaderless implementations.

The dependency-set abstraction can be implemented as shown in \cref{fig:deps-set-algo1,fig:deps-set-algo2}. Due to space constraints, our example focuses on satisfying the safety property only. A process $p$ announcing a command $c$ asks all the processes for the set of commands they have seen so far and that do not commute with $c$. Then, it suggests the union of the dependency sets received from a \emph{quorum} of processes. Quorums are sets of processes such that the intersection of any two
quorums is not empty. Since $p$ receives sets of commands seen by other processes, the implementation trivially ensures \ref{dep-valid}. 
Further, consider two suggestions $\opn{suggest}\left(c_1,\matthcal{D}_1\right)$ and $\opn{suggest}\left(c_2,\matthcal{D}_2\right)$ where $c_1$ and $c_2$ do not commute. There are two quorums $Q_1$ and $Q_2$ such that $\matthcal{D}_1$ was computed from $Q_1$ and $\matthcal{D}_2$ was computed from $Q_2$. By the definition of quorums, there is a process $q$ belonging to both $Q_1$ and $Q_2$. This process $q$ received either $c_1$ before $c_2$, in which case $c_1\in
\mathcal{D}_2$, or $c_2$ before $c_1$, in which case $c_2\in\mathcal{D}_1$. Therefore, property \ref{suggestion-agreement} holds.

The map-agreement algorithm can be implemented by any state machine replication algorithm, like MultiPaxos~\cite{lamport2001paxos}, where a command $c$ is uniquely associated with a position in the sequence of consensus instances. %used to implementing map-agreement by uniquely associating commands to MultiPaxos instance numbers.
However, MultiPaxos is too strong because we do not care about the ordering among instances.
Instead, one can uniquely associate one incarnation of MultiPaxos with each process, where each process is initially the leader of its MultiPaxos incarnation.  
We assume that a command is associated with a unique natural number $\opn{id}\left( c \right)$ and a process id $\opn{pid}\left( c \right)$, corresponding to the first process that received $c$.
A process receiving a proposal $\opn{propose}_p\left( c, \mathcal{D}) $  proposes $ c\mapsto\mathcal{D} $ in instance $\opn{id}\left( c \right)$ of the MultiPaxos incarnation of process $\opn{pid}\left( c \right)$. Upon a MultiPaxos decision $ c\mapsto\mathcal{D'} $, $p$ calls $\opn{commit}_p\left( c,\mathcal{D} \right)$.
Note that in a failure-free case, a command is committed in two message delays by its originator.

EPaxos improves upon our MultiPaxos scheme in two ways. First, EPaxos uses Fast Paxos\cite{Lamport05GeneralizeConsensus} incarnations instead of MultiPaxos incarnations. Second, EPaxos combines the dependency-set algorithm outlined above, but using fast quorums, with the first phase of Fast Paxos: the state of the dependency-set algorithm and of the Fast Paxos incarnations are merged so as to allow the reception of identical dependency sets from a fast quorum of processes
(after an $\opn{announce}_p\left( c \right)$) to act as a fast-path decision of Fast Paxos.

%A process receiving a proposal $ c\mapsto\mathcal{D} $ attaches a unique natural number $\opn{id}\left( c \right)$ as command identifier and its own id $\opn{pid}\left( c \right)$. proposes $ c\mapsto\mathcal{D} $ in instance $\opn{id}\left( c \right)$ of its MultiPaxos incarnation. Upon a MultiPaxos decision $ c\mapsto\mathcal{D'} $, $p$ commits $
%c\mapsto\mathcal{D'} $.

\printbibliography%
\begin{comment}
\appendix

@@

\end{comment}
\end{document}
